## Importing the required libraries

```{r}
library(ggplot2)
library(readxl)
library(dplyr)
```

### Read the Data set

```{r}
dt <- read.csv(file.choose())
```

```{r}
head(dt)
```

### Structure of the data

```{r}
str(dt)
```

### summary of the data

```{r}
summary(dt)
```

### Assign factors to character values

```{r}
dt$Date <- as.factor(dt$Date)
dt$Day <- as.factor(dt$Day)
dt$Time <- as.factor(dt$Time)
dt$CardType <- as.factor(dt$CardType)
dt$Gender <- as.factor(dt$Gender)
dt$BuyCategory <- as.factor(dt$BuyCategory)
dt$Region <- as.factor(dt$Region)
```

```{r}
#structure after factor the character columns
```

```{r}
str(dt)
```

### checking for null values in the data set

```{r}
colSums(is.na(dt))
```

```{r}
unique(dt$Time)
```

### Using Dplyr package

```{r}
data <- dt
```

### filter the data whose Buy category is High & analysis with the gender & Items Ordered column

```{r}
data %>% filter(BuyCategory == 'High') %>%
  select(Gender,ItemsOrdered,BuyCategory)%>%
  arrange(desc(ItemsOrdered))
```

### filter the Gender whose Buy category is High & Items Ordered is greater than 7

```{r}
data %>% filter(BuyCategory == 'High') %>%
  filter(Gender == 'Male' & ItemsOrdered >= 7)%>%
  select(Gender,ItemsOrdered,BuyCategory)%>%
  arrange(desc(ItemsOrdered))
```

### slice() gets row by index position

```{r}
data %>% slice(1,3,5)
```

### Get columns that end with given string:

```{r}
data %>% select(ends_with("Type")) %>% head()
```

### Get columns that match a string or regular expression:

```{r}
data %>% select(matches("Buy")) %>% head()
```

### Mutate() to add new variables to an existing data frame.

```{r}
data %>% mutate(TotalCost_all_item = ItemsOrdered * TotalCost) %>% head()
```

### summarize()

```{r}
data %>% summarize( Avg_TotalCost= mean(TotalCost))
```

# Data visualization

```{r}
hist(dt$TotalCost,xlab = "Total Cost", main ="Histogram of Total Cost",ylim = c(0,150))
```

```{r}
hist(dt$HighItem,main="Histogram of High_Item",xlab="High_item", ylim = c(0,200))
```

```{r}
hist(dt$ItemsOrdered,main="Histogram of ItemsOrdered",xlab="ItemsOrdered",ylim = c(0,200),xlim = c(0,12))
```

```{r}
plot(dt$ItemsOrdered,dt$TotalCost,,main = "Scatter plot ItemsOrdered vs TotalCost",xlab = "ItemsOrdered",ylab = "TotalCost", type= "p")
```

```{r}
ggplot(dt,aes(y = TotalCost, x = ItemsOrdered)) +geom_point()
```

```{r}
barplot(dt$ItemsOrdered,main = "ItemsOrdered bar graph",col = 'green',ylim = c(0,15))
```

```{r}
ggplot(dt,aes(y = Region, x = ItemsOrdered)) +geom_point()
```

```{r}
ggplot(dt) + geom_bar(mapping = aes(x=Region, color=Gender), fill='white', position='dodge')
```

```{r}
#table(), performs categorical tabulation of data with the variable and its frequency
counts <- table(dt$ItemsOrdered, dt$Gender)
barplot(counts, main = '',xlab="Number of Items",legend = rownames(counts),ylim = c(0,250))
```

```{r}
plot(density(dt$TotalCost), main='Total Cost Distribution')
```
